# Ultrasound

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
here::i_am("contents/ultrasound.qmd")
library(here)
library(dplyr)
library(tibble)
library(gt)
```

## Reference values

```{r us_nl, include=FALSE}
us_nl <- list(
  thyroid = "
* Size < 3 x 2 x 6 cm ?? (AP x TV x Long)
* Cervical LN
    * minimal axial diameter < 1 cm
    * oval shape (axis–to–long axis ratio [S/L] < 0.5)
",
  bladder = "
* < 0.8 cm (Collapse)
* < 0.4 cm (Distend)
",
  uterus = "
* Size < 5 x 4 x 8 cm (AP x TV x Long)
* ETT
    * < 1.5 cm
    * < 0.5 cm (post-menstrual)
",
  prostate = "
* < 25 ml (Normal)
* 25 - 40 ml (Prominent)
* \\> 40 ml (BPH)
",
  kidneys = "
Normal
* Long < 10 - 12 cm
* Cortical thickness > 1 cm

Abnormal
* Long < 8 cm (kidney disease)
",
  spleen = "
* < 12 cm (Normal)
* 12 - 15 cm (Mild splenomegaly)
* \\> 15 cm (Mark splenomegaly)
",
  veins = "
* Splenic v. < 11 cm
* SMV < 12 cm
* Portal v. < 13 cm
",
  aorta = "
Normal
* < 2.5 cm (Upper)
* < 2 cm (Lower)

Abnormal
* 2.5 - 3 cm (Ectasia)
* \\> 3 cm (Aneurysm)
",
  gb = "
* Long < 11 cm
* TV < 4 cm
* Wall thickness < 3 mm
",
  cbd = "
* < 0.6 cm (< 0.9 for post-cholecystectomy)
* Age \\> 60 yrs can dilate 0.1 cm / 10 yrs
"
)
```


```{r us_nl_tbl}
us_nl_tbl <- dplyr::tribble(
  ~Organs, ~`Reference Range`,
  # KUB
  "Kidneys", us_nl$kidneys,
  "Urinary Bladder (wall)", us_nl$bladder,
  "Residual Urine", "< 100 ml", 
  # Repro
  "Uterus", us_nl$uterus, 
  "Prostate", us_nl$prostate, 
  # Liver & GB
  "Gallbladder", us_nl$gb,
  "CBD", us_nl$cbd,
  # Vessel
  "Veins", us_nl$veins,
  "Aorta",  us_nl$aorta, 
  "Spleen", us_nl$spleen, 
  "Thyroid", us_nl$thyroid
) 

render_gt_us_nl_tbl <- function() {
  us_nl_tbl %>% 
  gt(rowname_col = "Organs") %>%
  fmt_markdown(columns = everything()) %>% 
  # Row Groups
  tab_row_group(
    label = "Vessel",
    rows = 8:9
  ) %>% 
  tab_row_group(
    label = "HBP",
    rows = 6:7
  ) %>% 
  tab_row_group(
    label = "Repro",
    rows = 4:5
  ) %>% 
  tab_row_group(
    label = "KUB", 
    rows = 1:3
    ) %>% 
  # FootNote
  tab_footnote("AP = Anteroposterior, TV = Transverse, Long = Longitudinal") %>% 
  # Style: Row Group
  tab_style(
    style = cell_text(weight = "bold", size = 20),
    locations = cells_row_groups()
  ) %>% 
  # Style: Column Labels
  tab_style(
    style = list(cell_text(weight = "bold", size = 20), 
                 cell_fill(color = "lightgrey")),
    locations = cells_column_labels()
  ) %>% 
  # Style: Stub
  tab_style(
    style = cell_text(weight = "bold" , size = 20),
    locations = cells_stub()
  ) %>% 
  # Style: Cell Body
  tab_style(
    style = cell_text(size = 20),
    locations = cells_body()
  ) %>% 
  tab_options(table.width = pct(90))
}

## Show for HTML only

render_gt_us_nl_tbl()
```




## Abdo Normal 

### Normal U/S WA


**FINDINGS:** 

**Liver:** Normal size and parenchymal echogenicity. No focal lesion. 

**Biliary system:** CBD size ___ mm. No intrahepatic ductal dilatation. 

**Gallbladder:** `[Well | Partially]`-distended gallbladder. No stone or mass.

**Spleen:** Normal in size. 

**Pancreas:** Visualized portions are unremarkable.

**Kidneys:** Normal size and parenchymal echogenicity of both kidneys. No stone, hydronephrosis or solid mass. 

**Urinary bladder:** No stone or mass. 

**Aorta:** Normal caliber. 

**Peritoneum:** No free fluid. 

(Choose one)

**Prostate gland:** Normal size, measuring ___ ml in volume. 

**Uterus:** Unremarkable.

**Ovaries:** `[Normal size of | Non-visualized]` both ovaries. No gross adnexal mass. 


**IMPRESSION:** [TODO]


### Normal U/S KUB

**FINDING:**

**Right kidney:** Normal size and parenchymal echogenicity, measuring about _ x _ x _ cm. No stone or hydronephrosis. 

**Left kidney:** Normal size and parenchymal echogenicity, measuring about _ x _ x _ cm. No stone or hydronephrosis.

**Bladder:** Well-distended urinary bladder without stone or mass. 

**Uterus:** Endometrial thickness measures about ___ mm. 

**Ovaries and adnexae:** Normal size of both ovaries.

**Others:** No free fluid. 

## Abdo Disease 

### Fatty Liver

**Liver:** Normal size and (diffusely) increased parenchymal echogenicity of the liver. 

- **(Mild):** mildly increased parenchymal echogenicity of the liver. 

- **(Moderate):** + with imparied visualization of hepatic vasculature,

- **(Severe):** + with imparied visualization of hepatic vasculature and diaphragm, 

- **(Focal fat sparing):** + A geographic hypoechoic lesion at periportal region, suggestive of focal fat sparing. 

... suggesting of `[mild | moderate | severe]` degree fatty liver disease.

+ No focal lesion. 

**IMPRESSION:** 

- `[Mild | Moderate | Severe]` degree fatty liver disease without focal lesion 
- Otherwise, unremarkable

### Gall Stone

**Gallbladder and biliary system:** A ???-cm gallstone. No gallbladder wall thickening or pericholecystic fluid. Common bile duct size ___ mm. No intrahepatic ductal dilatation. 

**IMPRESSION:** 

- A ???-cm gallstone without evidence of cholecystitis 
- Otherwise, unremarkable


### Dilate CBD

**Biliary system:** Dilated CBD measures about ___ mm 


- **(Not seen cause):** + without demonstrable cause of obstruction. No intrahepatic ductal dilatation. 


### Myoma & Cyst

**Uterus:** 

-   __ of a ? x ? x ?-cm `[intramural | submucosal | subserous]` myoma.

-   Endometrial thickness is measured about __ mm. 

**Ovaries and adnexae:** Normal size of ___ ovaries. A ???-cm cystic lesion at right adnexa, probably functional cyst or endometriotic cyst.

## Thyroid Template

### Normal Thyroid

**US OF THE THYROID** 

**FINDINGS:**

**Thyroid:** 

**Right lobe:** Normal size and homogeneous echogenicity. Size = _ x _ x _ cm. No mass. 

**Left lobe:** Normal size and homogeneous echogencity. Size = _ x _ x _ cm. No mass. 

**Isthmus:** Thickness = ___ mm. 

**Cervical lymph nodes:** No enlargement. 

**IMPRESSION:**

- [TODO]
- No cervical lymphadenopathy
